author: "Shalini Tyagi"

date: "14 December 2016"

output: html_document

---

## Overview

"The goal of this analysis is to predict the manner in which individuals performed an exercise. Six participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways. In the "classe" variable in the training set an "A" corresponds to the correct execution of the exercise, while the other 4 classes (B through E) correspond to common mistakes. By using data from accelerometers on the belt, forearm, arm, and dumbell we aim to predict which class the observation falls in."


```{r,echo=TRUE}
setwd("B:/shalini")
training<-read.csv("pml-training.csv")
testing<-read.csv("pml-testing.csv")
```

```{r,echo=TRUE}
library(caret)
library(rattle)
library(rpart)
library(randomForest)
```


## Cleaning up Training Data Set

```{r,echo=TRUE}
inTrain <- createDataPartition(training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]
myTesting <- training[-inTrain, ]
dim(myTraining); dim(myTesting)
```


```{r,echo=TRUE}
nzv <- nearZeroVar(myTraining, saveMetrics=TRUE)
myTraining <- myTraining[,nzv$nzv==FALSE]

nzv<- nearZeroVar(myTesting,saveMetrics=TRUE)
myTesting <- myTesting[,nzv$nzv==FALSE]
```
## Data spliting
 
```{r,echo=TRUE}
myTraining <- myTraining[c(-1)]
```

## Prediction Algorithms

```{r,echo=TRUE}
trainingV3 <- myTraining
for(i in 1:length(myTraining)) {
    if( sum( is.na( myTraining[, i] ) ) /nrow(myTraining) >= .7) {
        for(j in 1:length(trainingV3)) {
            if( length( grep(names(myTraining[i]), names(trainingV3)[j]) ) == 1)  {
                trainingV3 <- trainingV3[ , -j]
            }   
        } 
    }
}

# Set back to the original variable name
myTraining <- trainingV3
rm(trainingV3)
```


```{r,echo=TRUE}
clean1 <- colnames(myTraining)
clean2 <- colnames(myTraining[, -58])  # remove the classe column
myTesting <- myTesting[clean1]         # allow only variables in myTesting that are also in myTraining
testing <- testing[clean2]             # allow only variables in testing that are also in myTraining

dim(myTesting)
dim(testing)
```


```{r,echo=TRUE}
for (i in 1:length(testing) ) {
    for(j in 1:length(myTraining)) {
        if( length( grep(names(myTraining[i]), names(testing)[j]) ) == 1)  {
            class(testing[j]) <- class(myTraining[i])
        }      
    }      
}

# To get the same class between testing and myTraining
testing <- rbind(myTraining[2, -58] , testing)
testing <- testing[-1,]
```


## prediction

```{r,echo=TRUE}
library(rpart.plot)
set.seed(12345)
modFitA1 <- rpart(classe ~ ., data=myTraining, method="class")
fancyRpartPlot(modFitA1)
```

```{r,echo=TRUE}
predictionsA1 <- predict(modFitA1, myTesting, type = "class")
cmtree <- confusionMatrix(predictionsA1, myTesting$classe)
cmtree
```

```{r,echo=TRUE}
library(randomForest)
set.seed(12345)
modFitB1 <- randomForest(classe ~ ., data=myTraining)
predictionB1 <- predict(modFitB1, myTesting, type = "class")
cmrf <- confusionMatrix(predictionB1, myTesting$classe)
cmrf
```

## Prediction result on the test set

```{r,echo=TRUE}
predictionB2 <- predict(modFitB1, testing, type = "class")
predictionB2
```


